---
title: "Time Series Causal Analysis"
date: "`r format(Sys.Date(), '%Y-%m-%d')`" 
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: no
header-includes:
- \usepackage{booktabs}
- \usepackage{fancyhdr}
- \usepackage{tabu}
- \pagestyle{fancy}
- \fancyhead{}
params:
  dir: NA
  name: NA
  frequency: NA
  model: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\thispagestyle{fancy}

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(xts)
library(magrittr)
library(huxtable)
library(flextable)
library(tscausal)
library(rmarkdown)
x <- readRDS(paste0(params$dir,"/causal_tmp.rds"))
idate <- x$intervention_date
horizon <- NROW(x$post_actual)
```

# Model Report{-}

  * Series    : `r params$name`
  * Model Used: `r params$model`
  * Frequency : `r params$frequency`
  * Intervention Date: `r as.character(idate)`
  * Significance Level : `r 100*as.numeric(x$alpha)` \%
  * Hypothesis (H0): Intervention was not significant
  * Decision : `r ifelse(x$summary_table$p[1] <= x$alpha,"Reject","Fail to Reject")`


```{r,echo=FALSE,warning=FALSE,message=FALSE,ft.align="left"}
xx <- tscausal:::.table.print.tscausal(x, digits = 4)
tab <- as_hux(xx, add_colnames = TRUE) %>% add_rownames(colname = "Statistic", TRUE) %>% theme_grey() %>% map_align(by_cols("left","right","right")) %>% huxtable::set_caption("Causal Analysis") %>% huxtable::set_caption_pos("bottom") %>% huxtable::set_position("left")
tab$Statistic[1] <- "Statistic"
tab
```

```{r,echo=FALSE,warning=FALSE,message=FALSE,results='asis'}
p <- x$summary$p[1]
cat("Forecast Distribution tail-area probability p:   ", round(p, 5), "  \n")
cat("Forecast Distribution prob. of a causal effect:  ", round((1 - p) * 100, ifelse(p < 0.01, 5, ifelse(p < 0.05, 3, 0))), "\\%  \n")
```


```{r,echo=FALSE,warning=FALSE,message=FALSE}
plot(x)
```